---
title             : "Week 04 Reading and Lab2"
shorttitle        : "Reading and Lab2"

author: 
  - name          : "Pratik Chaudhari"
    affiliation   : ""
    corresponding : yes    # Define only one corresponding author
    address       : "326 Market Street, Harrisburg, PA 17101"
    email         : "pchaudhari@my.harrisburgu.edu"

affiliation:
  - id            : "247910"
    institution   : "Harrisburg University of Science and Technology"

authornote: |
  This document is a record of the summary of Fields book Chapter 3 and Lab2 scripts.

abstract: |

  
keywords          : "CRAN, menus, commands, workspace, packages, factor"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library(papaja)
library(knitr)
library(citr)
```


```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```




# R code output
#1,2
The if else provides the condition that when 2+2==5 then it returns FALSE. The x variable takes the value 8^13 and the y variable takes 15^9 and then we use less than (<), to check if x is less than or y or not and returns FALSE. 
```{r logicaloperators, echo=FALSE}
if(2+2==5){               #1
  print(TRUE)
} else {
  print(FALSE)
}

8^13< 15^9    #2
```


#3,4,5,6
We create a variable potato and assign a value 4 to it and then we calculate the square root of potato using sqrt() function and then check potato again to see that its value has not changed. We then reassign a new value to potato by multiplying original value by 2 using to get a new value of 8.
```{r variables, echo=FALSE}
potato<-4      
potato         
sqrt(potato)   
potato
potato<-potato*2   
potato
a<-"pratik"
a
b<- 3>5
b
c<-NA
c
```


#7,8,9,10,11,12
We create a numeric vector (num_vec), character vector (char_vec) and numeric vector (age) with three elements where each value of age corresponds to the name of each person. We use 'indexing by number to print first element of char_vec, "logical indexing" to print all ages greater (>) than 20 using. We create a data frame using the above three vectors and then print the age of one person using "indexing by name" which is 29.
```{r vectors, echo=F}
num_vec<-c(2,4,7)   
num_vec
char_vec<-c("aa", "cc", "ee")  
char_vec
age<-c(33,29,19)   
age

print(char_vec[3])  
print(age >20)   

df1 <- data.frame(num_vec,char_vec,age)   
df1
print(df1$age[df1$char_vec == "cc"])
```


#13,14,15,16,17,18,19,20,21,22,23
We load the airquality dataset using summary() and print the Wind variable from airquality using $. We print the third element of Wind which is 12.6. We create a new df "aq" with only first 10 cases from airquality and use logical indexing and subset() to print all days is aq where Ozone is greater than 20 and check the behaviour of NA values. We create a logical variable "TooWindy" inside aq to display TRUE when Wind is greater than 10 and FALSE when wind is less than 10.
Using length() function we determine number of observations in airquality. The mean of Wind in airquality is 9.958 and its standard deviation is 3.523. We plot the histogram of Ozone column and find out that it is not a normal distribution since mean and median are bith different.
```{r dataframes, echo=FALSE}
data(airquality)
summary(airquality) 
print(airquality$Wind)   
airquality$Wind[3]     

aq<-airquality[c(1:10), c("Ozone", "Solar.R", "Wind", "Temp", "Month", "Day")]  
aq
    
print(aq$Day[aq$Ozone > 20])     
subset(aq$Day, aq$Ozone>20)   

airquality$TooWindy<- ifelse(airquality$Wind >10, "TRUE", "FALSE")  

length(airquality)  

mean(airquality$Wind)   
table(aq$Temp)    

hist_ozone<-hist(airquality$Ozone)  
hist_ozone
```


#24,25
We create a simple sum() function to add x + 6 and add 6 to each value of "Temp" in airquality dataset.
```{r Functions, echo=FALSE}
x<-4
sum(x+6)    

print(airquality[, 4] +6)   
```


#26,27,28,29
Installing the packages.
\newline
# install.packages("ggplot2")
# install.packages("car")
# install.packages("ez")
```{r packages, echo=FALSE}
library(car)
```

\newpage

#30
We import Lab2.csv file and calculate dim(), str() and head() of the Lab2.
```{r files, echo=FALSE}
Lab2<-read.csv("/Users/one/Desktop/Principles-of-Analy-500/Assignments/HW-4/Lab2.csv", header=TRUE)
Lab2
dim(Lab2)
str(Lab2)
head(Lab2)
```


# R Packages
This document used `r cite_r("r-references.bib")` for R scripts and summary.


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
